
\chapter{Conclusions}
% "Koko hässäkän kiteytys mahdollisimman ytimekkäästi."

Requirement for offline support in single-page applications is not extraordinary today, and the need for it can be expected to grow in the future. Developers should keep that in mind when developing web applications and especially when designing software architecture of single-page applications.

Offline support can be implemented to an application with different levels of extent. The support can also be implemented to an already existing application, and in case the architecture is well designed, this can be done without extraordinary effort. Providing offline support may not require major changes to the server side of the system, but can be applied primarily to the code ran on the browser environment.

Knowing the specifics of the domain area helps in identifying the essential key activities within the application. Supporting only these activities during an Internet connection blackout is usually sufficient and will not sacrifice the user experience disturbingly much. This allows restricting of the scope of offline support, resulting in a fewer resources required on the development effort.

New features on the HTML standards -- such as \textit{application cache} and \textit{local storage} -- forms a great toolbox for implementing offline support for single-page applications. %% TODO tänne lisää

When looking from the user experience point of view, the technical methods introduced and evaluated in this thesis can also be beneficial in the development of single-page application in cases where there is no need for enabling usage of the application during a total Internet connection blackout. Applying these methods properly will result in a faster initialization times and smaller bandwidth usage than when using traditional approaches.

In cases where the status of the Internet connection can not be fully abstracted from the user's comprehension (only partial offline support is implemented), the concepts shown in the user interface must be considered thoroughly. For an average user the \textit{client-server model} on which Internet is built might not be distinct at all, and that can make motivating them to observe or care for the network status challenging. In scenarios like these, instructing the users about the context of the web applications in general might be the best alternative.
