
 \section{Software Quality in Futurice}
\label{sec:quality_futurice}

% http://www.futurice.com/wp-content/uploads/2014/02/Fact-Sheet-English2.pdf
Futurice is a lean service creation company founded in Finland. There are currently around 200 employees in four offices located in three countries. The employees of the company are passionate multi-talents in areas from technology, business consultancy and service design.~\cite{futuricefact}

Futurice does not have strict processes or common practices in quality assurance. The main idea present in all activities in the company is that smart people make smart decisions. With software quality, this means that the developers and other members of the project teams are able to decide their own approach to software quality. This freedom of choice leads to experimentation of multiple activities and practices, from which the good and bad experiences can be shared with others. Some practices of course become more common than others, but the use of those is not forced in any way. A common way of executing projects is to use methodologies from Lean and Agile software development.

In the core of Futurice, is communication and transparency. All the decisions and experiences are available for everyone and this brings the possibility to make decisions to every employee in the company. There are also informal communities built for areas of interests, such as mobile development and web development. These communities have recurring events where experiences are shared.

For the longer-term projects there is a separate team named Life Cycle Management team. Members from this team are present in a project from the beginning and they live through the product's life cycle participating in the development and maintenance of the product. This ensures that there is a solid amount of experience present in the team from the beginning of the project to the post-release phases.